---
title: "Income per capita and democracy score"
author: "Kyle Kehoe"
date: "2024-02-12"
output: html_document
---

```{r setup, echo=FALSE, message=F, warning=F, include=FALSE}

rm(list = ls())
setwd("C:\\Users\\kehky\\Downloads\\GDP_Democracy")
### Load the packages (all must have been installed)
     library(tidyverse)
    library(doBy)
    library(foreign)
    library(knitr)
    library(lmtest)
    library(readstata13)
    library(sandwich)
    library(stargazer)
    library(AER)
    library(gdata)
    library(wooldridge)
    library(openintro)


cse=function(reg) {
        rob=sqrt(diag(vcovHC(reg, type="HC1")))
        return(rob)
        }

```

```{r data, echo=FALSE, warning=FALSE, include=FALSE}
#Read the CSV file
incdemo <- read.csv("incdemo1995.csv", header=TRUE, sep=",")
```


```{r stargazer,echo=FALSE, warning=FALSE,comment=""}

stargazer(incdemo[c("log_gdppc", "dem_ind", "log_pop", "educ", "age_median")], type="text", title="Descriptive Statistics", summary.stat=c("n", "median", "mean", "sd","min", "max"), flip=FALSE)

```

From the descriptive statistics we can see there are 93 countries in this dataset.Based on the means and Medians and st.dev. we can infer that Log_GDP has moderate variability based on its st.dev, with only a slight left skew based on mean and median. Dem_ind appears to be left skewed and has a significant amount of variability in its data. Log_pop has a very mild right skew with significant deviation in the data. Educ is similar to the last very mild right skew with significant variability in the data. Finally Age_median once again has a moderate right skew with significant variability.
`

```{r images, echo=FALSE, warning=FALSE, message=FALSE,comment=""}
ggplot(incdemo, aes(y=log_gdppc, x=dem_ind)) + geom_point(col="blue") + geom_text(aes(label=country), hjust=1, vjust=1) +labs(title="Log Real GDP vs Democracy Index", x="Log Real GDP", y ="Democracy Index") + stat_smooth(method="lm", col= "red", se=FALSE)


```

We'll boot up a few regressions to look how some of these variables affect the regression and then determine which regression best models the affects democracy index on of real GDP per capita

```{r regression, echo=FALSE, warning=FALSE, message=FALSE, comment=""}

regr1<-lm(log_gdppc~dem_ind, data=incdemo)
regr2<-lm(log_gdppc~dem_ind+log_pop, data=incdemo)
regr3<-lm(log_gdppc~dem_ind+log_pop+educ, data=incdemo)
regr4<-lm(log_gdppc~dem_ind+log_pop+educ+age_median, data=incdemo)

# displays the regression results
stargazer(regr1, regr2, regr3,regr4, se=list(cse(regr1), cse(regr2), cse(regr3), cse(regr4)), 
        title="Real GDP on Democracy index", type="text", 
        star.cutoffs=NA, df=FALSE, digits=3, omit.stat=c( "f"))
```

Regression 1: a 1 unit increase in the democracy index is associated with a 2.11% increase in GDP
Regression 2: a 1 unit increase in the democracy index is associated with a 2.14% increase in GDP
Regression 3: a 1 unit increase in the democracy index is associated with a .60% increase in GDP
Regression 4: a 1 unit increase in the democracy index is associated with a .162% increase in GDP
These results are very interesting, it appears that there is string evidence of omitted variable bias, particularly with the large drop we see in anticipated GDP from Regression 2 to Regression 3 where it would be an omitted variable upward bias which means we have been over-anticipating the results on Real GDP for the democratic index. This leads me to believe that regression 3 and 4 are likely much more accurately predict the outcome democracy index has on GDP, especially with the large increase in R2 values where in regression 3 76% of GDP increase is accounted for by the democracy index while holding all 3 variables constant and Regression 4 has almost 83% accounted for.

Since Regression 4 seems to do the best job so far let's look at how 2 ficticious countries would compare, if we held all variables constant and only looked at the extremes of the democracy index what would the change on real gdp per capita be?
```{r prediction, echo=FALSE, warning=FALSE, message=FALSE}

dem0<-data.frame(dem_ind = 1,  log_pop = log(300000),  educ = 12,  age_median = 35  )
dem1<-data.frame(dem_ind =0,  log_pop = log(300000),  educ = 12,  age_median = 35  )

#predict(regr4, newdata=dem0)

#predict(regr4, newdata=dem1)

#Convert predictions of log real GDP per capita to real GDP per capita
print('predicted real gdp if Democracy ind=0, educ=12, log_pop=log(300K), median age=35')
exp(predict(regr4, newdata=dem0))
print('predicted real gdp if Democracy ind=1, educ=12, log_pop=log(300K), median age=35')
exp(predict(regr4, newdata=dem1))


```



If we hold the education, median age, and population constant and the only thing we change is the democracy index being full 1 or full 0 then real GDP per capita is expected to be $4468.4 higher. This is a significant increase of 17.6% over the Democracy index=0 country.

Finally let's do a linear hypothesis test to see if both population and median age should be in this regression.
```{r Linear hypothesis test, echo=FALSE, warning=FALSE, message=FALSE}
lht(regr4, c("log_pop=0", "age_median=0"), white.adjust="hc1")

```

the hypothesis test states that the p-value is approximately 0.00000002351 which is strong evidence that we should reject the null hypothesis and should keep at least one of the coefficients in the regression.

Overall from this entire exercise we can conclude that Real GDP per Capita is positively influenced by having a higher democracy index score. The strongest regression we had to model this was regression 4 which surprisingly predicted for 2 ficticious countries with a median age of 35, population of 300,000 people, education of 12 and the only differences being 1 country maxing out the Democracy index at 1 and another minimizing it's score at 0 there would be a 17.6% increase in real gdp per capita.

















